Hourly Assimilation of Different Sources of Observations Including Satellite Radiances in a Mesoscale Convective System Case During RELAMPAGO. Part I: Impact on the Analysis

### 2021-08-24 analysis/chapters/paper_ana_20181122.tex
Completo comentarios de Maru, Fig 13 ahora de 0 a 15 km

  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance. A multiphysics approach is also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations). We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files, AUT combining CONV and dense automatic surface weather station networks, SATWND, combining AUT with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI). We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study. The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
  \end{abstract}
   \begin{keyword} Regional Data Assimilation, Surface Obsevations, Satellite Observations\end{keyword}
 \end{frontmatter}

@@

DA methods can successfully be implemented by using observations with sufficient temporal and spatial resolution to characterize mesoscale atmospheric circulations.
However, surface observations, which directly measure the atmospheric state, have limited availability in South America.
Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products.
Radiance observations are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models.

The assimilation of conventional observations from surface stations has an important role in improving the initial conditions that lead to better forecasts.
For example, Ha and Snyder (2014) showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.

Given the sparsity of the local conventional observing network, it is important to explore the potential benefits of including additional sources of observations at a regional scale.
Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e.~AIRS and IASI) in a regional domain application.
Miyoshi and Kunii (2012) assimilated observations, atmospheric temperature, and humidity profile data from the AIRS and found improvement on the tracking of two typhoons.
Dillon et al. (2019) also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation.
Moreover, Bao et al. (2015) studied the impact on temperature and humidity forecast over the western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. Recently Zhu et al. (2019) studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels.
Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts (Gao et al., 2015).

@@

The model simulations for the case study were performed using the 3.9.1 version of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), Skamarock et al. (2018)) model.
The resolution was 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa.
The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25\(^{\circ}\) horizontal resolution and 6-hour temporal resolution, National Centers for Environmental Prediction, National Weather Service, NOAA, U.S. Department of Commerce (2015)).
The domain covers the area indicated in Figure \ref{fig:dominio} to capture the development of the MCS during the simulated period.

The analyses are generated using the Local Ensemble Transform Kalman Filter (LETKF) implementation (V1.3, Hunt et al. (2007)) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; Hu et al. (2018)).
A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within \(\pm\) 30 minutes of the analysis time are assimilated.
Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals.
For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al. (2006)) was used as an observation operator to calculate model-simulated brightness temperatures.

@@

In addition to random perturbations at the lateral boundaries, a multi-physics scheme is used to better represent the uncertainty in model formulation within the DA system. We use 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain--Fritsch (Kain, 2004), Grell--Freitas (Grell and Freitas, 2013), and Betts--Miller--Janjic (Janjić, 1994)) and 3 planetary boundary layer schemes (Yonsei University Scheme (Hong, Noh, et al., 2006), Mellor--Yamada--Janjic Scheme (Janjić, 1994), and Mellor--Yamada Nakanishi Niino (Nakanishi and Niino, 2009)). All ensemble members used the same land-surface model (Noah-MP, Chen and Dudhia (2001)), microphysics (WRF single-moment 6--class scheme (Hong, Kim, et al., 2006)), and radiation processes (RRTMG shortwave and longwave scheme (Iacono et al., 2008)) parameterizations.

\begin{figure}
\includegraphics[width=1\linewidth]{../figures/dominio-1} \caption{a) The horizontal domain used on the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles). b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8 and purple square are radiosondes launched outside the IOP missions. The topography in meters is also shown (filled contours).}\label{fig:dominio}
\end{figure}

@@

The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS). We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at NCEP that consist of surface observations from conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircraft. The orange triangles in Figure \ref{fig:dominio}a, are the surface stations and ship locations included in this experiment. The orange triangles in Figure \ref{fig:dominio} a), are the surface station locations included in this experiment. The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region.

@@

\caption{\label{tab:table-obs}Characteristics of the assimilated observations: The code for each observation type and it source, the available variables, the observation error and the gross check thresholds used.}
\centering
\fontsize{6}{8}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5em}>{\raggedright\arraybackslash}p{4.5em}>{\raggedright\arraybackslash}p{5em}>{\raggedright\arraybackslash}p{7em}>{\raggedright\arraybackslash}p{7em}}
\toprule
Code & Platform & Variable & Error & Gross check\\
\midrule
 &  & Pressure & 1-1.6 $hPa^*$ & 3.6 $hPa$\\

@@

Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2). Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, the number of satellite observations varied significantly among cycles. This is evident in Figure 3 a) which shows the number of radiance observations assimilated at each analysis cycle as a function of pressure. In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 \% of the total amount of assimilated radiances in our experiment. The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM. The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere. The channels whose maximum weighting function is located in the stratosphere were rejected due to the low model top (50 hPa) used in the experiments.

@@

In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves for the \(\beta_i\) that provides the best fit between the simulation and the observations. The coefficients initialized from the 18 UTC Nov 11, 2018, GFS system coefficients were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section. Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each cycle.

@@

\caption{\label{tab:table-rad}List of the available sensors over several platforms, the number of accepted channels for the assimilation and the percentage of assimilated observations calculated over all radiance observations and all cycles.}
\centering
\fontsize{7}{9}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedleft}X>{\raggedright}X}
\toprule
Sensor & Plataform & \makecell[r]{Assimilated\\channels} & \makecell[l]{Percentage\\over total}\\
\midrule
AIRS & AQUA & 118 & 39.37 \%\\
\cmidrule{1-4}
 & AQUA & 7 & 1.13 \%\\
\cmidrule{2-4}
 & N19 & 13 & 1.55 \%\\
\cmidrule{2-4}
 & N15 & 13 & 3.93 \%\\
\cmidrule{2-4}
 & N18 & 13 & 3.17 \%\\
\cmidrule{2-4}
\multirow[t]{-5}{*}{\raggedright\arraybackslash AMSUA} & METOP-A & 13 & 1.3 \%\\
\cmidrule{1-4}
 & N20 & 21 & 1.16 \%\\
\cmidrule{2-4}
\multirow[t]{-2}{*}{\raggedright\arraybackslash ATMS} & NPP & 21 & 0.86 \%\\
\cmidrule{1-4}
 & N19 & 14 & 1.79 \%\\
\cmidrule{2-4}
\multirow[t]{-2}{*}{\raggedright\arraybackslash HIRS4} & METOP-A & 14 & 1.37 \%\\
\cmidrule{1-4}
 & METOP-A & 174 & 40.07 \%\\
\cmidrule{2-4}
\multirow[t]{-2}{*}{\raggedright\arraybackslash IASI} & METOP-B & 174 & 2.94 \%\\
\cmidrule{1-4}
 & N19 & 5 & 0.44 \%\\
\cmidrule{2-4}
 & METOP-A & 5 & 0.46 \%\\
\cmidrule{2-4}
\multirow[t]{-3}{*}{\raggedright\arraybackslash MHS} & METOP-B & 5 & 0.44 \%\\
\bottomrule
\end{tabu}
\end{table}

@@

\hypertarget{experimental-design}{%
\subsection{Experimental design}\label{experimental-design}}

@@

The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \ref{fig:obs-horizontal}. The larger number of assimilated observations over the center and east of the domain corresponds to the ASWS observations. In this region, we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet. In Figure \ref{fig:obs-cycle}a the number of assimilated observations over time is shown. Local maxima at 12 and 00 UTC found mainly in CONV is produced by the operational soundings. The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage. The maxima after 12 and 00 UTC in RAD correspond to the contribution of the multispectral sensors. The vertical distribution of the mean number of observations per cycle (Figure \ref{fig:obs-cycle}b) shows a maximum in low levels due to the ASWS observations. Satellite-derived winds available in this particular case are maximum at the upper troposphere (between 500-250 hPa). Above 850 hPa, most of the observations correspond to radiance observations.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/home/paola.corrales/mesoda/analysis/figures/analysis_cycle} \caption{Diagram of the analysis cycles between Nov 20th, 18 UTC and Nov 23 12 UTC plus and spin up period of 6 hours. The zoomed section shows the hourly assimilation hour that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours. The two Intensive Observation Periods (IOP) missions from the RELAMPAGO Field Campaign are shown.}\label{fig:cycle}
\end{figure}

@@

where \(O\) and \(X\) stand for independent observations and the analysis respectively.

For the comparison of the first-guess' precipitation with the precipitation estimates, we compute a probabilistic Fraction Skill Score (FSS) for different sized sampling areas and thresholds, following Maldonado et al. (2021) and adaptations from Roberts (2008).

@@

We performed all analysis in this paper using the R programming language (R Core Team, 2020), using data.table (Dowle and Srinivasan, 2020) and metR (Campitelli, 2020) packages.
All graphics are made using ggplot2 (Wickham, 2009) and the paper was rendered using knitr and rmarkdown (Allaire et al., 2020; Xie, 2015). All the experiments were performed at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne (Computational and Information Systems Laboratory, 2019).

@@

Figure \ref{fig:rcrv-profile} shows the mean and standard deviation of the RCRV for the upper-air observations. Figure \ref{fig:rcrv-profile}a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT). Both ADPUPA and AIRCFR show a generally good agreement between the ensemble spread and the observation error. As sounding observations and their associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread. AIRCAR presents an irregular profile with \(sd RCRV\) values that suggest that the error for this type of observation is overestimated. ADPUPA and AIRCAR present a \(mean RCRV\) profile near zero on middle and upper levels but ADPUPA observations show a cold bias with respect to the model at low levels, a characteristic already studied in Ruiz et al. (2010) and Dillon et al. (n.d.).

For satellite-derived winds (Figure \ref{fig:rcrv-profile}c) in low levels, where there are not many observations available, the profiles of \(mean RCRV\) and \(sd RCRV\) shows a larger departure from the expected behavior with a negative bias and an overestimation of the observation error. Wind estimations derived from water vapor channels are abundant over 500 hPa where their bias is close to zero. The only exception are the EUMETSAT observations which contribute very little in the region.

@@

The impacts on the wind components are shown in Figure Figure \ref{fig:UV-diff} along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations. The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis. There is a diurnal cycle in the impact of surface weather stations on the meridional velocity (Figure \ref{fig:UV-diff}d) with a stronger reduction of the northerly wind during night hours. This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment. After 18 UTC Nov, 22 the opposite effect is observed when the MCS is moving through the domain. After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels. During Nov 22 and 23 the impact of assimilating ASWS produces an increase of northerly wind in upper levels. This could be a consequence of a stronger MCS with an increased polar side upper level outflow. Satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, however, the systematic impact is overall smaller than the one produced by assimilating data from ASWS.

The assimilation of radiances produced a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure \ref{fig:UV-diff}c and f). For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of \(1 ms^{-1}\), opposite to what is generated by the assimilation of ASWS observations. On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind. When we look at the meridional wind field at 200 hPa at different times, we observed that the outflow from the MCS is even more intense than in the other experiments while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in Figure \ref{fig:UV-diff}f.~

To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the low level northerly flow, precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles). At that moment the first convective cells were developing on the southern region of the domain along the cold front. Figure \ref{fig:summary-fields}a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours). We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations. The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place. AUT and SATWND experiments are very similar, with values of precipitable water over 55 \(kgm^{-2}\) north of 30\(^{\circ}\)S and a similar vertical distribution of specific humidity (not shown). RAD has lower precipitable water content than AUT and SATWND but higher than CONV. The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating ASWS -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compare to CONV.

>>>>

  We used the GSI-4DLETKF data assimilation package to produce analyses assimilating observations every hour with 10-km horizontal grid spacing and a 60-members ensemble initialized from the deterministic GFS run adding random perturbations with climatological covariance. A multiphysics approach was also used to represent model errors, using different physics configurations (a combination of PBL and convection parameterizations). We conducted four assimilation experiments using different sets of observations: CONV, consisting of conventional observations from NCEP's prepBUFR files, AUT combining CONV and dense automatic surface weather station networks, SATWND, combining AUT with satellite-derived winds and RAD, including SATWND and satellite radiances from different microwave and infrared sensors (AMSU, HIRS, MHS, ATMS, AIRS, and IASI). We found that the assimilation of observations with high temporal and spatial frequency generate an important impact on the PBL, primarily on the precipitable water content, that leads to the development of deep convection and heavy precipitation closer to the observed in this case study. The assimilation of radiance observations produces a better development of the convection mainly during the mature state of the MCS leading to an increase in the accumulated precipitation.
  \end{abstract}
   \begin{keyword} Regional Data Assimilation, Surface Obsevations, Satellite Observations\end{keyword}
 \end{frontmatter}

@@

DA methods can successfully be implemented by using observations with sufficient temporal and spatial resolution to characterize mesoscale atmospheric circulations.
The assimilation of conventional observations from surface stations has an important role in improving the initial conditions that lead to better forecasts.
For example, Ha and Snyder (2014) showed that the assimilation of temperature and dew point temperature systematically improved the structure of the simulated planetary boundary layer that subsequently resulted in a better precipitation forecast over the U.S.
However, surface observations, which directly measure the atmospheric state, have limited availability in South America. Given the sparsity of the local conventional observing network, it is important to explore the potential benefits of including additional sources of observations at a regional scale.
Other types of observations are being assimilated with relatively high spatial and temporal resolution, such as satellite radiance observations or retrieval products.
Previous work has shown a positive impact at incorporating temperature and humidity profiles derived from the multispectral infrared sensors (i.e.~AIRS and IASI) in a regional domain application.
Miyoshi and Kunii (2012) assimilated observations, atmospheric temperature, and humidity profile data from the AIRS and found improvement on the tracking of two typhoons near Japan.
Dillon et al. (2019) also assimilated retrievals from the AIRS sensor and reported improvements on the thermodynamic variables and the regional circulation over Southern South America.

However radiance observations are indirect measurements and are more difficult to incorporate into DA systems as they require complex forward models. Moreover, Bao et al. (2015) studied the impact on temperature and humidity forecast over the western USA while assimilating microwave and infrared radiance data and found a reduction in the temperature bias at low and mid-levels as a result of the microwave observations but an opposite effect for infrared data. Recently Zhu et al. (2019) studied the impact of assimilating satellite radiance data within a frequently updated regional system and showed an improvement for all variables, in particular for relative humidity at upper levels.
Satellite-derived winds from water vapor channels and cloud drift motion vectors also have an overall positive impact on analysis and short-range forecasts (Gao et al., 2015).

@@

The model simulations for the case study are performed using the 3.9.1 version of the non-hydrostatic Advanced Research version of the Weather Research and Forecasting (WRF-ARW (V3.9), Skamarock et al. (2018)) model.
The resolution is 10 km (150 x 200 grid points) on the horizontal and 37 levels on the vertical with the top of the atmosphere at 50 hPa.
The initial and boundary conditions are provided by the Global Forecasting System (GFS) analysis (0.25\(^{\circ}\) horizontal resolution and 6-hour temporal resolution, National Centers for Environmental Prediction, National Weather Service, NOAA, U.S. Department of Commerce (2015)).
The domain covers the area indicated in Figure \ref{fig:dominio} to capture the development of the MCS during the simulated period.

The analyses are generated using the Local Ensemble Transform Kalman Filter (LETKF) implementation (V1.3, Hunt et al. (2007)) part of the Gridpoint Statistical Interpolation analysis system (GSI V3.8; Hu et al. (2018)).
A rapid update cycle approach is implemented with hourly analysis and a centered assimilation window, meaning that all the observations within \(\pm\) 30 minutes of the analysis time are assimilated.
Observations are assimilated in a 4D approach by comparing them with the corresponding first guess state at 10-minute intervals.
For radiance observations, the Community Radiative Transfer Model version 2.3 (CRTM; Han et al. (2006)) is used as an observation operator to calculate model-simulated brightness temperatures.

@@

In addition to random perturbations at the lateral boundaries, a multi-physics scheme is used to better represent the uncertainty in model formulation within the DA system. We use 9 different model configurations consisting of the combination of 3 moist convection schemes (Kain--Fritsch (Kain, 2004), Grell--Freitas (Grell and Freitas, 2013), and Betts--Miller--Janjic (Janjić, 1994)) and 3 planetary boundary layer schemes (Yonsei University Scheme (Hong, Noh, et al., 2006), Mellor--Yamada--Janjic Scheme (Janjić, 1994), and Mellor--Yamada Nakanishi Niino (Nakanishi and Niino, 2009)). All ensemble members use the same land-surface model (Noah-MP, Chen and Dudhia (2001)), microphysics (WRF single-moment 6--class scheme (Hong, Kim, et al., 2006)), and radiation processes (RRTMG shortwave and longwave scheme (Iacono et al., 2008)) parameterizations.

\begin{figure}
\includegraphics[width=1\linewidth]{../figures/dominio-1} \caption{a) The horizontal domain used on the simulations (black box), the inner domain used for the experiment comparison (red box), the region shown in b) (light blue box) and the locations of Automatic Surface Weather Stations (ASWS, green squares) and Conventional Surface Weather Stations (CSWS, orange triangles). b) Locations of radiosonde launches during the experiment, green dots corresponds to radiosondes launched during IOP 7, orange triangles are radiosondes launched during IOP 8 and purple square are radiosondes launched outside the IOP missions. The topography in meters is also shown (shaded).}\label{fig:dominio}
\end{figure}

@@

The conventional and satellite-derived wind observations used are part of the Global Data Assimilation System (GDAS). We assimilated conventional observations included in the Binary Universal Form for Representation of Meteorological Data (PREPBUFR) files generated at National Centers for Environmental Prediction (NCEP) that consist of surface observations from 117 conventional surface weather stations (CSWS) and ships and upper-air observations from radiosondes and aircraft. The orange triangles in Figure \ref{fig:dominio} a), are the surface station locations included in this experiment. The frequency of these observations varied between 1 hour for surface stations and 12/24 hours for most radiosondes in the region. Wind surface observations over oceans (ASCATW) come from scaterometters and are also part of the PREPBUFR files.

@@

\caption{\label{tab:table-obs}Characteristics of the assimilated observations: The code for each observation type and its source, the available variables, the observation error and the gross check thresholds used.}
\centering
\fontsize{6}{8}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5em}>{\raggedright\arraybackslash}p{4.5em}>{\raggedright\arraybackslash}p{5em}>{\raggedright\arraybackslash}p{7em}>{\raggedright\arraybackslash}p{7em}}
\toprule
Code & Platform & Variable & Error & Gross check\\
\midrule
 &  & Pressure & 1-1.6 $hPa^*$ & 3.6 $hPa$\\

@@

Satellite radiances available through the GDAS data stream, consisting of infrared and microwave observations are used in this study. This includes the Advanced Microwave Sounding Unit - A (AMSU-A), Advanced Technology Microwave Sounder (ATMS), Microwave Humidity Sounder (MHS), High-resolution Infrared Radiation Sounder (HIRS/4), and 2 multispectral sensors, Atmospheric Infrared Sounder (AIRS) and Infrared Atmospheric Sounding Interferometer (IASI) over several satellite platforms (see Table 2). Since the regional domain is located in the mid-latitudes and the satellite platforms of interest are on polar orbits, each sensor scans the area only twice a day with a spatial coverage depending on the satellite swath. For this reason, the number of satellite observations varied significantly among cycles. This is evident in Figure 3 a) which shows the number of observations assimilated for RAD experiment (see Section \ref{exp}) at each analysis cycle. In particular, the multispectral sensors provided between 100 and 10.000 observations for every scan every 12 hours, contributing 82 \% of the total amount of assimilated radiances in our experiment. The vertical location of each radiance observation was estimated as the model level at which its weighting function was maximized as calculated by CRTM. The multispectral sensors have good vertical coverage and are able to sense from the lower troposphere and up to the lower stratosphere. The channels whose maximum weighting function is located in the stratosphere were rejected due to the low model top (50 hPa) used in the experiments.

@@

In the GSI system, the \(\beta_i\) coefficients are trained using a variational estimation method which solves the \(\beta_i\) that provides the best fit between the simulation and the observations. The coefficients were initialized at 18 UTC Nov 11 with the GFS system coefficients, and were trained during one week of continuous assimilation using the same setup as in the experiments described in the next section. Moreover, the assimilation system was configured to use a constant background error variance of 0.01 to avoid large adjustments in the estimated coefficients at each cycle.

@@

\caption{\label{tab:table-rad}List of the available sensors over several platforms, the number of accepted channels for the assimilation and the percentage of assimilated observations calculated over all radiance observations and all cycles.}
\centering
\fontsize{7}{9}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedleft}X>{\raggedright}X}
\toprule
Sensor & Platform & \makecell[r]{Assimilated\\channels} & \makecell[l]{Percentage\\over total}\\
\midrule
AIRS & AQUA & 118 & 39.37 \%\\
\cmidrule{1-4}
 & AQUA & 7 & 1.13 \%\\
\cmidrule{2-4}
 & N19 & 13 & 1.55 \%\\
\cmidrule{2-4}
 & N15 & 13 & 3.93 \%\\
\cmidrule{2-4}
 & N18 & 13 & 3.17 \%\\
\cmidrule{2-4}
\multirow[t]{-5}{*}{\raggedright\arraybackslash AMSUA} & METOP-A & 13 & 1.3 \%\\
\cmidrule{1-4}
 & N20 & 21 & 1.16 \%\\
\cmidrule{2-4}
\multirow[t]{-2}{*}{\raggedright\arraybackslash ATMS} & NPP & 21 & 0.86 \%\\
\cmidrule{1-4}
 & N19 & 14 & 1.79 \%\\
\cmidrule{2-4}
\multirow[t]{-2}{*}{\raggedright\arraybackslash HIRS4} & METOP-A & 14 & 1.37 \%\\
\cmidrule{1-4}
 & METOP-A & 174 & 40.07 \%\\
\cmidrule{2-4}
\multirow[t]{-2}{*}{\raggedright\arraybackslash IASI} & METOP-B & 174 & 2.94 \%\\
\cmidrule{1-4}
 & N19 & 5 & 0.44 \%\\
\cmidrule{2-4}
 & METOP-A & 5 & 0.46 \%\\
\cmidrule{2-4}
\multirow[t]{-3}{*}{\raggedright\arraybackslash MHS} & METOP-B & 5 & 0.44 \%\\
\bottomrule
\end{tabu}
\end{table}

@@

\hypertarget{exp}{%
\subsection{Experimental design}\label{exp}}

@@

The horizontal distribution of the average number of assimilated observations per cycle in each experiment is shown in Figure \ref{fig:obs-horizontal}. The larger number of assimilated observations over the center and east of the domain corresponds to the ASWS observations. In this region, we observed high values of precipitation during Nov 22 and is associated with the inflow of moisture by the South American low level jet (Vera et al., 2006). In Figure \ref{fig:obs-cycle}a the number of assimilated observations over time is shown. Local maxima at 12 and 00 UTC found mainly in CONV is produced by the operational soundings. The strong variability in the number of radiance observations per cycle is also noticeable and depends on the satellite coverage. The maxima after 12 and 00 UTC in RAD correspond to the contribution of the multispectral sensors. The vertical distribution of the mean number of observations per cycle (Figure \ref{fig:obs-cycle}b) shows a maximum in low levels due to the ASWS observations. Satellite-derived winds available in this particular case are maximum at the upper troposphere (between 500-250 hPa). Above 850 hPa, most of the observations correspond to radiance observations.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/home/paola.corrales/mesoda/analysis/figures/analysis_cycle} \caption{Diagram of the analysis cycles between Nov 20, 18 UTC and Nov 23 12 UTC plus and spin up period of 6 hours. The zoomed section shows the hourly assimilation that is performed within a one-hour centered window and new boundary conditions from GFS every 6 hours. The two IOPs missions from the RELAMPAGO Field Campaign are shown.}\label{fig:cycle}
\end{figure}

@@

where \(O\) and \(X\) stand for independent observations and the simulations respectively.

For the comparison of the first-guess' precipitation with the precipitation estimates, we compute a probabilistic Fractions Skill Score (FSS) for different sized sampling areas and thresholds, following Maldonado et al. (2021) and adaptations from Roberts (2008):

@@

We performed all the experiments at the National Center for Atmospheric Research (NCAR) supercomputer Cheyenne (Computational and Information Systems Laboratory, 2019). All the analysis in this paper were performed using the R programming language (R Core Team, 2020), using data.table (Dowle and Srinivasan, 2020) and metR (Campitelli, 2020) packages.
All graphics are made using ggplot2 (Wickham, 2009) and the paper was rendered using knitr and rmarkdown (Allaire et al., 2019; Xie, 2015).

@@

Figure \ref{fig:rcrv-profile} shows the mean and standard deviation of the RCRV for the upper-air observations. Figure \ref{fig:rcrv-profile}a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT). Both ADPUPA and AIRCFT show a generally good agreement between the ensemble spread and the observation error. As sounding observations and their associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread. AIRCAR presents an irregular profile with \(sd RCRV\) values that suggest that the error for this type of observation is overestimated. ADPUPA and AIRCAR present a \(mean RCRV\) profile near zero on middle and upper levels but ADPUPA observations show a cold bias with respect to the model at low levels, a characteristic already studied in Ruiz et al. (2010) and Dillon et al. (n.d.).

For satellite-derived winds (Figure \ref{fig:rcrv-profile}c) in low levels, where there are not many observations available, the profiles of \(mean RCRV\) and \(sd RCRV\) show a larger departure from the expected behavior with a negative bias and an overestimation of the observation error. Wind estimations derived from water vapor channels are abundant over 500 hPa where their bias is close to zero. The only exception are the EUMETSAT observations which contribute very little in the region.

@@

The impacts on the wind components are shown in Figure \ref{fig:UV-diff} along with the corresponding averaged wind component in the experiment with the largest number of assimilated observations. The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis. There is a diurnal cycle in the impact of surface weather stations on the meridional velocity (Figure \ref{fig:UV-diff}d) with a stronger reduction of the northerly wind during night hours. This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment. After 18 UTC Nov, 22 the opposite effect is observed when the MCS is moving through the domain. After the initiation of the convective cells, the systematic impact on the wind field is larger at mid and upper levels. During Nov 22 and 23 the impact of assimilating ASWS produces an increase of northerly wind in upper levels. This could be a consequence of a stronger MCS with an increased polar side upper level outflow. Although satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, the systematic impact is overall smaller than the one produced by assimilating data from ASWS.

The assimilation of radiances produces a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure \ref{fig:UV-diff}c and f). For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of \(1 ms^{-1}\), opposite to what is generated by the assimilation of ASWS observations. On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind. When we look at the meridional wind field at 200 hPa at different times, we observe that the outflow from the MCS is even more intense than in the other experiments, while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in Figure \ref{fig:UV-diff}f.~

To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the low level northerly flow, precipitable water, low level temperature, and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles). At that moment the first convective cells were developing on the southern region of the domain along the cold front. Figure \ref{fig:summary-fields}a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours). We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations. The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place. AUT and SATWND experiments are very similar, with values of precipitable water over 55 \(kgm^{-2}\) north of 30\(^{\circ}\)S and a similar vertical distribution of specific humidity (not shown). RAD has lower precipitable water content than AUT and SATWND but higher than CONV. The distribution of moisture at low levels in RAD seems to be the result of the combination of the moistening effect of assimilating ASWS -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the center of the domain compared to CONV.

The potential temperature on the PBL (Figure \ref{fig:summary-fields}b) resembles the characteristics observed on the temperature profiles in Figure \ref{fig:TQ-diff}a-c.~On average the PBL in AUT and SATWND is colder than in CONV while RAD shows a warmer PBL than AUT due to the assimilation of radiance observations. Figure \ref{fig:summary-fields}c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 \(km\) wind shear. The values of MCAPE in CONV do not exceed 2000 \(JKg^{-1}\) while the rest of the experiments show maximum MCAPE over 4000 \(JKg^{-1}\). MCAPE in the RAD experiment is lower compared to AUT or SATWND. This is consistent with less humidity in the PBL with respect to these experiments and may be partially compensated by a slightly warmer PBL in the RAD experiment. The 0-6 km wind shear is more intense in AUT, SATWND, and RAD reaching values over 15 \(ms^{-1}\) at the southern tip of the region with positive MCAPE values. Moreover, in this same region, these experiments show larger MCAPE values than CONV. Note that wind shears over 15 \(ms^{-1}\) have been associated with the development of more intense and organized MCSs (Chen et al., 2015) and also with conditions favorable for supercells (Markowsky and Richardson, 2010).
